\bf\mftitle

### 2011-02-23 doc/jInfer/autoeditor/autoeditor.tex
Documentation: AutoEditor modified according to the review remarks. Not finished yet.

\section{Introduction}
This is an implementation of a \jmodule{AutoEditor}. Using JUNG library, it provides an API to display and user interactively modify automata, thus the process of inference can be easily made user interactive. For more information on JUNG library, see \cite{jung}.\\

UML diagrams shown in this document are simplified to keep them synoptical. Simplification involves removing not important class members which are not mentioned in the text, removing members of not jInfer classes and truncating string of inheritance which are not important for this document.\ For example, \code{VisualizationViewer} class from JUNG library has many methods and does not extends \code{JPanel} class directly, though in an UML diagram it has not any methods and extends \code{JPanel} directly, because it is sufficient for understanding of this document.

@@

\subsection{Layout}
\code{Layout} is a JUNG interface responsible primarily of representation of automaton and positions of its states. JUNG library provides several implementation of \code{Layout} interface. However, none of them is convenient for automatic automaton displaying, \jmodule{AutoEditor} provides two additional implementations. \code{Layout} by \emph{Julie Vyhnanovska}, used in her master thesis and \code{Layout} which is using external \emph{Graphviz} software. TODO odkazy na juliu, graphviz a JUNG v predchadzajucej vete.\\

@@

\subsubsection{Vyhnanovska Layout} \label{subsubsection:vyhnanovskaLayout}
As mentioned above, this Layout was implemented by \emph{Julie Vyhnanovska} as a part of her master thesis. TODO link? It positions automaton states to a square grid. This \code{Layout} gives good results for relatively small automata (about 10 states of less) but for larger ones, the results are quite disarranged and confused.\\

Source codes reside in package\newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts.vyhnanovska}.

@@

Automaton is transformed to the \emph{dot} representation and the \emph{dot} binary is ran with this representation on input. It processes the automaton and writes positions of its states to the standard output. Than the output is parsed and an instances of \code{Layout} is created using these positions.\\

Source codes reside in package\newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts.graphviz}.

\subsubsection{LayoutHelperFactory} \label{subsubsection:LayoutHelperFactory}
In project properties, it is possible to select a \code{Layout} to be used to display automata. \code{LayoutHelperFactory} is class, providing just one static method, responsible for creating instances of \code{Layout}s according to a selection in project properties.\\

This method has following signature.
\begin{verbatim}
public static <T> Layout<State<T>, Step<T>>
createUserLayout(Automaton<T> automaton, Transformer<Step<T>, String> edgeLabelTransformer)
\end{verbatim}
The first argument is a automaton to create a layout from. The second is transformer to transform an instance of automaton edge to its string representation, required by the \code{Graphviz Layout}.\\

Source codes resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts}.

@@

\code{Layout}s can be implemented using the modular system. To create a new implementation of \code{Layout} interface, it is needed to create a new class implementing \code{LayoutFactory} interface (package \code{package cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts}) and annotate it by the following code.

@@

System of modules is described in documentation of Architecture, see \cite{archdoc}.

@@

\code {Visualizer} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} package extends JUNG \linebreak \code{VisualizationViewer} class thus it provides all its methods and adds support for saving contained automaton to an image file. Responsible methods are \code{saveImage()} and \code{getSupportedImageFormatNames()}. However, to save an image of automaton it is not necessary to call this methods directly. \jmodule{AutoEditor} GUI contains button to save an image of displayed automaton. For information on how to to this, see \ref{subsection:GUI}.\\

Constructor has one argument, instance of \code{Layout} interface created from an automaton, typically by \linebreak \code{LayoutHelperFactory} (see \ref{subsubsection:LayoutHelperFactory}).

@@

By default, instance of \code{PluggableVisualizer} is constructed with two plugins enabled. They are \linebreak \code{ScalingGraphMousePlugin}, providing zooming, and \code{TranslatingGraphMousePlugin}, providing translating the displayed automaton in the x and y direction. In the most cases, these plugins are useful but if they are not wanted they can be removed using methods \code{getGraphMousePlugins()} and \code{removeGraphMousePlugin()}.\\

@@

\code{AbstractComponent} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component} package is a representation of GUI panel containing an instance of \code{Visualizer} class for some automaton. It is inherited from \code{JPanel} class thus provides JPanel's method and behaviour. In addition, it provides the following methods.

\begin{itemize}
	\item \code{setVisualizer()} - Setter of \code{Visualizer}.
	\item \code{getVisualizer()} - Getter of \code{Visualizer}.
	\item \code{waitForGuiDone()} - Suspends its thread until method \code{guiDone} is called on this instance. Do not call this method directly, it is called by \jmodule{AutoEditor}. For more information, see \ref{subsubsection:AbstractComponent user-interactivity support}.
	\item \code{guiDone()} - Wakes up this instance from a suspended state. For detailed description, see \ref{subsubsection:AbstractComponent user-interactivity support}.
	\item \code{guiInterrupt()} - Called when \jmodule{AutoEditor}'s tab is closed to propagate information about terminating of inference to a caller of \jmodule{AutoEditor}. There is no need to called this method directly.
	\item \code{guiInterrupted()} - Checks if \jmodule{AutoEditor} GUI was terminated by interrupt or regularly. Also, there is no need to call this method directly, it is called by \jmodule{AutoEditor}. For details, see \ref{subsection:GUI}.
\end{itemize}

@@

Programmer implementing an extension of \code{AbstractComponent} is not forced to place the \code{Visualizer} on its own. It is just needed to create \code{JPanel} and define \code{getAutomatonDrawPanel()} method to return this \code{JPanel}. \jmodule{AutoEditor} will take care of placing and displaying the \code{Visualizer} in the \code{JPanel}.\\

@@

If some kind of user interactivity is desired, \code{AbstractComponent} is a right place to implement it.\\

To display the component in GUI and wait for some user action, method \code{waitForGuiDone()} is used. After displaying the component, calling of this method suspend running thread thus code execution of a caller module is stopped at the place of this call. However, GUI is ran in another thread, user is able to interact with the panel (component).\\*
Do not call method \code{waitForGuiDone()} directly. It is called by \jmodule{AutoEditor} when displaying the component by \jmodule{AutoEditor} \emph{API} method named \code{drawComponentAndWaitForGUI()}. For more information on \jmodule{AutoEditor} \emph{API}, see \ref{subsection:API}.\\

Method important for a programmer extending \code{AbstractComponent} is named \code{guiDone()}. This method wakes up the thread suspended in \code{waitForGuiDone()} method and the programmer is responsible for calling it. Typically, it is called upon some user action like button click, vertex pick or other GUI event.\\*
After calling of \code{guiDone()} method, code execution of the caller module is resumed and holding instance of the \code{AbstractComponent} it is able to retrieve results of user interaction, saved in its state.\\

@@

\begin{itemize}
	\item \code{drawComponentAsync()} - Displays given \code{AbstractComponent} asynchronously in a GUI thread and immediately returns. Use this method to just display automaton, without any user interaction and without waiting for any external event. This method does not support these.
	\item \code{drawComponentAndWaitForGUI()} - Displays given \code{AbstractComponent} in a GUI thread and a caller thread is suspended until  \code{guiDone()} method of \code{AbstractComponent()} is called. This method can wait for GUI events thus is convenient for user interaction. How to achieve it is described in detail in \ref{subsubsection:AbstractComponent user-interactivity support}.
	\item \code{closeTab()} - Closes \jmodule{AutoEditor}'s GUI tab and interrupts inference, if running.
\end{itemize}

@@

This section describes classes derived from the base classes used in \jmodule{Two Step Simplifier} module to display automata. This classes may also serve as examples of extending the base classes.

@@

Source codes of \code{StatePickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component}, \code{StatePickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and \linebreak \code{VertexPickingGraphMousePlugin} in \newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.graphmouseplugins}.\\

Example of usage follows.
\begin{verbatim}
    final Transformer<Step<Regexp<T>>, String>
      transformer = new Transformer<Step<Regexp<T>>, String>() {

@@

    State<Regexp<T>> removeState;
    final StatePickingComponent<Regexp<T>> component = new StatePickingComponent<Regexp<T>>();
    final Layout<State<T>, Step<T>> layout = LayoutHelperFactory.createUserLayout(automaton,
                                                                                  transformer);
    final StatePickingVisualizer<Regexp<T>>
      visualizer = new StatePickingVisualizer<Regexp<T>>(layout,
                                                         transformer,
                                                         component,
                                                         automaton.getSuperInitialState(),
                                                         automaton.getSuperFinalState());
    component.setVisualizer(visualizer);

@@

\code{StatesPickingComponent} is almost the same as \code{StatePickingComponent} but it contains two extra GUI controls. 'Continue' button and 'Don't ask me anymore to select states to be merger' checkbox.\
The button is supposed to be pushed when user picked all desired states and it will cause calling of component's \code{guiDone()} method (see \ref{subsubsection:AbstractComponent}).\
State of the checkbox can be retrieved by the calling code and the caller is supposed to stop showing this component in the current inference process.

Methods of \code{StatesPickingComponent} are the following.
\begin{itemize}
	\item \code{shallAskUser()} - Retrieves state of the checkbox.
	\item \code{getPickedStates()} - After displaying the component using \code{drawComponentAndWaitForGUI()} \emph{API} method (see \ref{subsection:API}), this method retrieves the list of user picked automaton states. In case that none of automaton states was picked, it returns an empty list.
	\item \code{setLabel()} - Sets text of a component label. The label can be used to communicate some information to user, for example instructions.
\end{itemize}

Source codes of \code{StatesPickingComponent} resides in package \newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component}, \newline \code{StatesPickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and \newline \code{VerticesPickingGraphMousePlugin} in \newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.graphmouseplugins}.\\

@@

    List<State<T>> mergeLst;
    final Transformer<Step<T>, String> transformer = new Transformer<Step<T>, String>() {

@@

    Boolean selectTwo = false;
    do {
      final Layout<State<T>, Step<T>>
        layout = LayoutHelperFactory.createUserLayout(inputAutomaton, transformer);
      final StatesPickingVisualizer<T>
        visualizer = new StatesPickingVisualizer<T>(layout, transformer);

      final StatesPickingComponent<T> panel = new StatesPickingComponent<T>();
      panel.setVisualizer(visualizer);
      if (selectTwo) {
        panel.setLabel("Automaton for element <" + elementName
          + ">. Please select states to be merged and click continue. Select at least 2 states.");
      } else {
        panel.setLabel("Automaton for element <" + elementName
          + ">. Please select states to be merged and click continue.");
      }

@@

As shown at figure \ref{figure_gui}, \jmodule{AutoEditor}'s panel is placed in a tab in the main NetBeans window. It consists of two buttons, horizontal line below them and below the line, there is panel to place an extension of \code{AbstractComponent} (see \ref{subsubsection:AbstractComponent}). Class representing this panel is named \code{AutoEditorTopComponent} and resides in \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.topcomponent} package.\\

Buttons have labels 'Save as image' and 'Show settings'. Pushing the first one will raise a dialog box to save currently displayed automaton to an image file. Set of supported image formats depends on installed JRE.\
The second one will open \jmodule{AutoEditor} tab in NetBeans options. For description of the settings, see \ref{subsection:Settings}.

>>>>

\section{Introduction}
This is an implementation of a an automaton editor. Using JUNG library, it provides an API to display and modify automata in an interactive mode. For more information on JUNG library, see \cite{jung}.\\

UML diagrams shown in this document are simplified to keep them readable. Simplification involves removing not important class members which are not mentioned in the text, removing members of non-jInfer classes and truncating string of inheritance which are not important for this document.\ For example, \code{VisualizationViewer} class from JUNG library has many methods and does not extends \code{JPanel} class directly, though in an UML diagram it has not any methods and extends \code{JPanel} directly, because it is sufficient for understanding of this document.

@@

\subsection{Layout}
\code{Layout} is a JUNG interface responsible primarily of representation of an automaton and positions of its states. JUNG library provides several implementation of \code{Layout} interface. However, because none of them is convenient for automatic automaton displaying, \jmodule{AutoEditor} provides two additional implementations. \code{Layout} by \emph{Julie Vyhnanovska}, used in her master thesis and \code{Layout} which is using external \emph{Graphviz} software. TODO odkazy na juliu, graphviz a JUNG v predchadzajucej vete.\\

@@

\subsubsection{Vyhnanovska Layout} \label{subsubsection:vyhnanovskaLayout}
As mentioned above, this Layout was implemented by \emph{Julie Vyhnanovska} as a part of her master thesis. TODO link? It positions automaton states on a square grid. This \code{Layout} gives good results for relatively small automata (about 10 states of less) but for larger ones, the results are quite disarranged and confusing.\\

Source code resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts.}\\*\code{vyhnanovska}.

@@

Automaton is transformed to the \emph{dot} representation and the \emph{dot} binary is invoked with this representation on input. It processes the automaton and writes positions of its states to the standard output. The output is then parsed and an instance of \code{Layout} is created using these positions.\\

Source code resides in package\newline \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts.graphviz}.

\subsubsection{LayoutHelperFactory} \label{subsubsection:LayoutHelperFactory}
In project properties, it is possible to select a \code{Layout} to be used to display automata. \code{LayoutHelperFactory} is a class providing just one static method responsible for creating instances of \code{Layout}s according to a selection in project properties.\\

This method has the following signature.
\begin{verbatim}
public static <T> Layout<State<T>, Step<T>>
createUserLayout(Automaton<T> automaton, Transformer<Step<T>, String> edgeLabelTransformer)
\end{verbatim}
The first argument is an automaton to create layout from. The second is transformer to transform an instance of automaton edge to its string representation, required by the \code{Graphviz Layout}.\\

Source code resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts}.

@@

New \code{Layout}s can be implemented using the modular system. For information on system of modules, see \cite{archdoc}. To create a new implementation of \code{Layout} interface, it is necessary to create a new class implementing \code{LayoutFactory} interface (package \code{package cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.layouts}) and annotate it by the following code.

@@

@@

\code {Visualizer} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} package extends JUNG \linebreak \code{VisualizationViewer} class thus it provides all its methods and adds support for saving contained automaton to an image file. Relevant methods are \code{saveImage()} and \code{getSupportedImageFormatNames()}. However, to save an image of automaton it is not necessary to call this methods directly. \jmodule{AutoEditor} GUI contains button to save an image of displayed automaton. For information on how to to this, see \ref{subsection:GUI}.\\

Constructor has one argument, an instance of \code{Layout} interface created from an automaton, typically by \linebreak \code{LayoutHelperFactory} (see \ref{subsubsection:LayoutHelperFactory}).

@@

By default, instance of \code{PluggableVisualizer} is constructed with two plugins enabled. They are \linebreak \code{ScalingGraphMousePlugin}, providing zooming, and \code{TranslatingGraphMousePlugin}, providing translating the displayed automaton in the x and y direction. In the most cases these plugins are useful, otherwise they can be removed using methods \code{getGraphMousePlugins()} and \code{removeGraphMousePlugin()}.\\

@@

\code{AbstractComponent} class in \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.component} package is a representation of GUI panel containing an instance of \code{Visualizer} class for some automaton. It is inherited from \code{JPanel} class, thus provides JPanel's method and behaviour. In addition, it provides the following methods.

\begin{itemize}
	\item \code{setVisualizer()} - Setter of \code{Visualizer}.
	\item \code{getVisualizer()} - Getter of \code{Visualizer}.
	\item \code{waitForGuiDone()} - Suspends its thread until method \code{guiDone} is called on this instance. Do not call this method directly, it is called by \jmodule{AutoEditor}. For more information, see \ref{subsubsection:AbstractComponent user-interactivity support}.
	\item \code{guiDone()} - Wakes up this instance from a suspended state. For detailed description, see \ref{subsubsection:AbstractComponent user-interactivity support}.
	\item \code{guiInterrupt()} - Called when \jmodule{AutoEditor}'s tab is closed to propagate information about terminating of inference to a caller of \jmodule{AutoEditor}. There is no need to called this method directly.
	\item \code{guiInterrupted()} - Checks if \jmodule{AutoEditor} GUI was terminated by interrupt or regularly. Again, there is no need to call this method directly, it is called by \jmodule{AutoEditor}. For details, see \ref{subsection:GUI}.
\end{itemize}

@@

Programmer implementing an extension of \code{AbstractComponent} is not forced to place the \code{Visualizer} on his own. It is just needed to create \code{JPanel} and define \code{getAutomatonDrawPanel()} method to return this \code{JPanel}. \jmodule{AutoEditor} will take care of placing and displaying the \code{Visualizer} in the \code{JPanel}.\\

@@

If some kind of user interactivity is desired, \code{AbstractComponent} is the right place to implement it.\\

To display the component in GUI and wait for some user action, method \code{waitForGuiDone()} is used. After displaying the component, calling this method suspends the running thread, thus code execution of a caller module is stopped at the place of this call. However, GUI is ran in another thread, user is able to interact with the panel (component).\\*
Do not call method \code{waitForGuiDone()} directly. It is called by \jmodule{AutoEditor} when displaying the component by \jmodule{AutoEditor} \emph{API} method named \code{drawComponentAndWaitForGUI()}. For more information on \jmodule{AutoEditor} \emph{API}, see \ref{subsection:API}.\\

Method important for a programmer extending \code{AbstractComponent} is called \code{guiDone()}. This method wakes up the thread suspended in \code{waitForGuiDone()} method and the programmer is responsible for calling it. Typically, it is called upon some user action like button click, vertex pick or other GUI event.\\*
After calling of \code{guiDone()} method, code execution of the caller module is resumed and holding instance of the \code{AbstractComponent} it is able to retrieve results of user interaction, saved in its state.\\

@@

\begin{itemize}
	\item \code{drawComponentAsync()} - Displays given \code{AbstractComponent} asynchronously in a GUI thread and immediately returns. Use this method to just display automaton, without any user interaction and without waiting for any external event. This method does not support it.
	\item \code{drawComponentAndWaitForGUI()} - Displays given \code{AbstractComponent} in a GUI thread, while the caller thread is suspended until  \code{guiDone()} method of \code{AbstractComponent()} is called. This method can wait for GUI events and thus is convenient for user interaction. How to achieve this is described in detail in \ref{subsubsection:AbstractComponent user-interactivity support}.
	\item \code{closeTab()} - Closes \jmodule{AutoEditor}'s GUI tab and interrupts inference, if running.
\end{itemize}

@@

This section describes classes derived from the base classes used in \jmodule{Two Step Simplifier} module to display automata. This classes may also serve as examples of base classes extensions.

@@

Source codes of \code{StatePickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.}\linebreak\code{component}, \code{StatePickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and\linebreak \code{VertexPickingGraphMousePlugin} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.}\linebreak\code{graphmouseplugins}.\\

Example of usage follows.
\begin{verbatim}
    Transformer<Step<Regexp<T>>, String>
      transformer = new Transformer<Step<Regexp<T>>, String>() {

@@

    State<Regexp<T>> removeState;
    StatePickingComponent<Regexp<T>> component = new StatePickingComponent<Regexp<T>>();
    Layout<State<T>, Step<T>> layout = LayoutHelperFactory.createUserLayout(automaton, transformer);
    StatePickingVisualizer<Regexp<T>>
      visualizer = new StatePickingVisualizer<Regexp<T>>(layout,
                                                         transformer,
                                                         component,
                                                         automaton.getSuperInitialState(),
                                                         automaton.getSuperFinalState());
    component.setVisualizer(visualizer);

@@

\code{StatesPickingComponent} is almost the same as \code{StatePickingComponent} but it contains two extra GUI controls. 'Continue' button and 'Don't ask me anymore to select states to be merger' checkbox.\
The button is supposed to be clicked when user picked all desired states and it will cause calling of component's \code{guiDone()} method (see \ref{subsubsection:AbstractComponent}).\
State of the checkbox can be retrieved by the calling code and the caller is supposed to stop showing this component in the current inference process.

Methods of \code{StatesPickingComponent} are the following.
\begin{itemize}
	\item \code{shallAskUser()} - Retrieves state of the checkbox.
	\item \code{getPickedStates()} - After displaying the component using \code{drawComponentAndWaitForGUI()} \emph{API} method (see \ref{subsection:API}), this method retrieves the list of user picked automaton states. In the case that none of automaton states was picked, it returns an empty list.
	\item \code{setLabel()} - Sets text of the component label. The label can be used to communicate some information to user, for example instructions.
\end{itemize}

Source codes of \code{StatesPickingComponent} resides in package \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.}\linebreak\code{component}, \code{StatesPickingVisualizer} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer} and\linebreak \code{VerticesPickingGraphMousePlugin} in \code{cz.cuni.mff.ksi.jinfer.autoeditor.automatonvisualizer.}\linebreak\code{graphmouseplugins}.\\

@@

    List<State<T>> mergeLst;
    Transformer<Step<T>, String> transformer = new Transformer<Step<T>, String>() {

@@

    Boolean selectTwo = false;
    do {
      Layout<State<T>, Step<T>>
        layout = LayoutHelperFactory.createUserLayout(inputAutomaton, transformer);
      StatesPickingVisualizer<T> visualizer = new StatesPickingVisualizer<T>(layout, transformer);

      StatesPickingComponent<T> panel = new StatesPickingComponent<T>();
      panel.setVisualizer(visualizer);
      if (selectTwo) {
        panel.setLabel("Automaton for element <" + elementName
          + ">. Please select states to be merged and click continue. Select at least 2 states.");
      } else {
        panel.setLabel("Automaton for element <" + elementName
          + ">. Please select states to be merged and click continue.");
      }

@@

As shown at figure \ref{figure_gui}, \jmodule{AutoEditor}'s panel is placed in a tab in the editor window. It consists of two buttons, horizontal line below them and and a panel to place an extension of \code{AbstractComponent} (see \ref{subsubsection:AbstractComponent}). Class representing this panel is named \code{AutoEditorTopComponent} and resides in \code{cz.cuni.mff.ksi.jinfer.autoeditor.gui.topcomponent} package.\\

Buttons have labels 'Save as Image' and 'Show Settings'. Pushing the first one will raise a dialog box to save currently displayed automaton to an image file. Set of supported image formats depends on installed JRE.\
The second one will open \jmodule{AutoEditor} tab in NetBeans options. For description of the settings, see \ref{subsection:Settings}.
